---
title: "Chapter 6: Mathematics of Networks"
output: html_notebook
---
```{r setup, echo = F}
source("../classes/Node.R")
source("../classes/Edge.R")
source("../classes/SimpleNetwork.R")
```
# 6.1: Networks and their representations

This chapter starts with a basic definition of a network or graph as a collection of *nodes*, which are joined together by *edges*.

Using this simple definition, I start off by defining S4 classes for each of these items.

, here are the items that should be implemented for a new class definition:

Core
* definition
* validity method
* constructor function
* "show" method

API
* getters
* setters

I've defined a node class, which only has one slot corresponding to its name.
```{r define node}
getClass("Node")
```

With this class, I can perform some tests.
```{r test Node}
# Creating a test node
test_node <- Node()

# Testing the getter
name(test_node)

# Testing the setter
name(test_node) <- "node1"
name(test_node)
```

Next, I define an edge, which has two attributes: a source node, and a target node. Because this item is slightly more complicated, I also define a setter to create a new edge from two provided nodes.

```{r define edge}
getClass("Edge")
```

```{r test Edge}
# Creating two test nodes
node1 <- Node("node1")
node2 <- Node("node2")
node3 <- Node("node3")
node4 <- Node("node4")

# Creating some test edges
edge1 <- new("Edge", source_node = node1, target_node = node2)
edge2 <- new("Edge", source_node = node2, target_node = node1)
edge3 <- new("Edge", source_node = node3, target_node = node4)

# Testing the getters
source_node(edge1)
target_node(edge1)

# Testing the setters
source_node(edge1) <- node3
target_node(edge1) <- node4

source_node(edge1)
target_node(edge1)

# Testing equality
all.equal(edge1, edge1)
all.equal(edge1, edge2)
all.equal(edge1, edge3)
```


```{r define network}
getClass("SimpleNetwork")
```

```{r create SimpleNetwork}
test_network <- new("SimpleNetwork", node_list = list(node1, node2, node3, node4), edge_list = list(edge1, edge3))
```

```{r test Bad Network}
# Testing validator function
bad_network <- new("SimpleNetwork", node_list = list(node1, node2, "bad_node", node4), edge_list = list(edge1, "bad_edge"))
```

```{r test SimpleNetwork}
node5 <- new("Node", name = "node5")
test_network <- add_node(test_network, node5)



# Testing edge add and delete
node99 <- new("Node", name = "node99")
node98 <- new("Node", name = "node98")
edge99 <- new("Edge", source_node = node99, target_node = node98)
get_edge_index(test_network, edge99)
test_network <- add_edge(test_network, edge99)

test_network <- remove_edge(test_network, edge99)
```


